---
title: "lab_06"
author: "Abigail Russ"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* [install Tabula](https://tabula.technology/)
* run existing code as directed (look for **Task**).
* write code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes in the lab_06 folder and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Get Our PDF

We'll be working with the [Maryland Statistical Handbook](https://planning.maryland.gov/MSDC/Documents/md-statistical-handbook.pdf) from the state Department of Planning. It's not tiny (44MB), and you'll want to download it to a place you'll remember (like your Downloads folder).

**Task** Download the file from the above link and save it to your Downloads folder.

## Setup Tabula

**Task** Start Tabula and load the PDF, then extract a table

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the Statistical Handbook file in your Downloads folder and click "open", and then click the "Import button" in Tabula. This will take a minute or two.

Let's go to page 30 of the PDF, Table 2A, "International Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". This is _net migration_, meaning it shows the total number of people moving to Maryland and its regions/jurisdictions _minus_ the total number of people leaving the same, divided into 12-month periods from July to July. In Tabula, draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV (it should be called `tabula-md-statistical-handbook.csv` by default) to your lab_06/data folder.

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)):

**Task** Write code to read the CSV file into a dataframe called `international_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores.

```{r}
international_migration <- read_csv("data/tabula-md-statistical-handbook.csv", skip=2, col_names=c("State/Jurisdiction","7/1/2010-7/1/2011", "7/1/2011-7/1/2012", "7/1/2012-7/1/2013","7/1/2013-7/1/2014", "7/1/2014-7/1/2015","7/1/2015-7/1/2016","7/1/2016-7/1/2017","7/1/2017-7/1/2018","7/1/2018-7/1/2019","4/1/2010-7/1/2019"))
```

Add a column for the type of migration ("international") and populate it:

**Task** Write the code to add a `migration_type` column to the data and give it the value of "international"

```{r}
international_migration %>% 
mutate(migration_type = "international")
```

## Answer questions

**Q1** Which region and county/city _below_ the state level accounted for the largest percentage of international migration for the entire time period? You'll need to add and populate a column representing percent of total using `mutate`.
**A1** Over the entire time period the Suburban Washington Region accounted for the most international migration at 62.84% and Montgomery County was the county with the largest international migration at 36.79%.

```{r}
international_migration %>% 
  mutate(percent_total=(`4/1/2010-7/1/2019`)/max(`4/1/2010-7/1/2019`)*100) %>% 
  arrange(desc(percent_total)) %>% 
  select(`State/Jurisdiction`, percent_total)

```

**Q2** Write a sentence or two that describes the data you produced in A1. Try to convey the most important idea.
**A2** The data I found breaks down the net international migration to the state of Maryland by region and by county. It is very important to note that some of the information could be double counted if you're not focusing because people can be counted twice, once in the region and once in the county. For example, someone who migrated to Montgomery County could also be counted as someone who migrated to the Suburban Washington Area so you must make sure you're separating the region and the county. In addition, I notice that there is a lot of migration to urban areas like the Baltimore Region and the Suburban Washington Area and the migration numbers drop significantly. Rural Maryland has much lower international migration. 

**Q3** Which region & jurisdiction had the largest percentage change (either negative or positive) for international migration between July 2018 and July 2017? The formula for percentage change is easy to remember: (New-Old)/Old.
**A3** Calvert County had the largest percent change. The county's international migration decreased by 76.92%. The region with the largest percent change was the Southern Maryland Region. 
## our numbers are just the change so what column is the old and new -- what is the old population. we only know the change in population which is the net migration. 

the end of july 2017 to the end of july 2018 old is 16-17

```{r}
international_migration %>% 
  mutate(percent_change = (((`7/1/2017-7/1/2018`)-(`7/1/2016-7/1/2017`))/(`7/1/2016-7/1/2017`))*100) %>%
  arrange(percent_change) %>% 
  select(`State/Jurisdiction`, percent_change)

```

**Q4** What's your best guess as to why these changes occurred, and in those places in particular?
**A4** I would guess that these decreases in international migration occurred particularly in Calvert County because it is an overwhelmingly white area and July 2017 was just a few months after Trump's travel ban. It could be possible that international migration decreased in rural areas like Southern Maryland and Calvert County in favor of more urban areas that could have more favorable views of international immigration. In addition, Calvert County is one of the richest country in America so it's possible that it is too expensive to live in the area. Calvert County is in the Southern Maryland Region so it makes sense that these decreases are occurring together. 

## Back to Tabula

**Task** Extract second table from Tabula

Let's go to page 31 of the PDF, Table 2B, "Domestic Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". In Tabula, hit the "Clear All Selections" button and then draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV to your lab_06/data folder. (and rename it to `tabula-md-statistical-handbook-domestic.csv`).

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner:

**Task** Write code to read the CSV file into a dataframe called `domestic_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores. Your work on the international file should serve as a guide.

```{r}
domestic_migration <- read_csv("data/tabula-md-statistical-handbook-domestic.csv", skip=2, col_names=c("State/Jurisdiction","7/1/2010-7/1/2011", "7/1/2011-7/1/2012", "7/1/2012-7/1/2013","7/1/2013-7/1/2014", "7/1/2014-7/1/2015","7/1/2015-7/1/2016","7/1/2016-7/1/2017","7/1/2017-7/1/2018","7/1/2018-7/1/2019","7/1/2010-7/1/2019"))
```

Add a column for the type of migration ("domestic") and populate it:

**Task** Write the code to add a `migration_type` column to the data and populate it with the value of "domestic"

```{r}
domestic_migration %>% 
  mutate(migration_type = "domestic")
```

## Answer questions
**Q5** Which Maryland individual jurisdiction saw the largest net decrease in domestic migration overall?
**A5** The Suburban Washington Region had the largest net decrease in domestic migration overall, but Baltimore City had the largest net decrease of domestic migration of the cities/counties. 

```{r}
domestic_migration %>% 
  select(`State/Jurisdiction`, `7/1/2010-7/1/2019`) %>% 
  arrange(`7/1/2010-7/1/2019`)

```

**Q6** How many regions & jurisdictions had net positive migration for July 2017, July 2018 and July 2019 (not just individually but all three together)?
**A6** There were 14 regions and jurisdictions in Maryland with positive net migrations for July 2017, July 2018 and July 2019.  

##filter each of them to be more than 0 and then add
```{r}
domestic_migration %>% 
  filter(`7/1/2016-7/1/2017` > 0 & `7/1/2017-7/1/2018` >0 & `7/1/2018-7/1/2019`>0) %>% 
  select(`State/Jurisdiction`, `7/1/2016-7/1/2017`, `7/1/2017-7/1/2018`, `7/1/2018-7/1/2019`)

```

**Q7** How would you describe this data? Is there a county or region that stands out, and why?
**A7** After analyzing the data, I think it is interesting that the regions with positive domestic migration are more rural. This is the opposite of what we found with international data. For example, Frederick County has the highest domestic migration rate over the three years listed. That is probably the closest county to a major metropolitan area. In addition, Frederick County borders Montgomery County which has a negative net domestic migration. I wonder if people may be leaving the major metropolitan areas (which have negative net values) towards to more rural parts of Maryland (like from Montgomery County to Frederick County). 
