---
title: "lab_04"
author: "Abigail Russ"
date: "2/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
maryland_ppp <- read_rds("lab_04.rds")
# add code here to accomplish the task
```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the pre-lab. There's a particular set of steps you'll need to perform to make that work properly, so take your time and check your work.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
naics_codes <- read_csv("data/naics_codes.csv")
# add code here to accomplish the task

```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.

```{r}
maryland_ppp <- maryland_ppp %>% left_join(naics_codes, by="naics_code") %>%
  mutate(zip5 = str_sub(zip, start=1L, end=5L)) 
```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one.

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character. IMPORTANT: you can see what the column names represent by checking the data documentation, which can be found here: https://geodata.md.gov/imap/rest/services/Demographics/MD_CensusData/FeatureServer/1. You will need to refer to this to answer the questions below.

```{r}
maryland_zcta <- read_csv("data/maryland_zcta.csv") %>% 
  mutate(ZCTA5N= as.character(ZCTA5N))
```

Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
 maryland_ppp <- maryland_ppp %>% left_join(maryland_zcta, by=c("zip5" = "ZCTA5N"))
```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?
**A1** The top 5 industries with the most PPP applications are 722511	(Full-Service Restaurants),621111 (Offices of Physicians (except Mental Health Specialists)), 238220	(Plumbing, Heating, and Air-Conditioning Contractors), 541110 (Offices of Lawyers), 236220 (Commercial and Institutional Building Construction). 

```{r}
maryland_ppp %>% 
  group_by(naics_code) %>% 
  summarise(total_amount = sum(amount),industry=first(title)) %>% 
  arrange(desc(total_amount))

```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** In order from most to least, the top 5 industries with PPP loans that were over 150,000 were Full-Service Restaurants (with 1488 loans), Offices of Physicians (except Mental Health Specialists)	with 580 loans, Plumbing, Heating, and Air-Conditioning Contractors	with 458 PPP loans, Offices of Dentists that got 419 loans and finally Commercial and Institutional Building Construction which received 402 PPP loans over 150,000. 

```{r}
maryland_ppp %>% 
  filter(amount > 150000) %>% 
  group_by(naics_code,title) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) 
```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** The top zip codes with at least 20 percent non-Hispanic Asian population with the largest sum of PPP loans is 20850 (Rockville)	with $474525058	and 3186 loans. In second place was North Potomac's 20878	with $151646470	and 2188 loans. Finally there's Ellicott City's 21043 with $72404137 across 1385 PPP loans. 

```{r}
maryland_ppp %>% 
  filter(PNHA>20) %>% 
  group_by(zip5) %>% 
  summarize(total_amount = sum(amount), n=n()) %>% 
  arrange(desc(total_amount))
  
```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4** Rockville (20850) had the largest sum of loans where the reported race of the applicant was Asian. The percentage of the non-Hispanic Asian population in 20850 is 23.5% and there were 240 PPP loans. 

```{r}
maryland_ppp %>% 
  filter(race=="Asian") %>% 
  group_by(zip5,PNHA) %>% 
  summarize(total_amount = sum(amount), n=n()) %>% 
  arrange(desc(total_amount))

```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants?
**A5** With such a significant percentage of Asian identifying people in 20850, only 7.5% of the loans in the zip code may have gone to Asian applicants because Asian people in that specific zip code may not own small businesses that qualify for PPP loans or there may be more Asian people who chose not to identify their race when filling out the form, which would skew our data.


**Q6** Check to see if any Maryland zip codes that have loans where the PPP data lists the applicant as being Asian BUT are located in a zip code that does not appear to have Asian population according to the ZCTA data. How many loans do they have and what is the total amount? Take a look at these zip codes - what could explain this?
**A6** There are 3 zip codes that say they have a 0% Asian population but have applicants self identifying as Asian when applying for PPP loans.  20701	has 4 loans worth 75170.50 dollars, 21031 had 6 loans worth 6819563.00 dollars and 21240 had 3 loans worth 358674.00 dollars. This could be explained by a rounding error in the percent Asians in a zip code. They probably rounded down to 0 if they had a marginal percentage of Asian people in a zip code. 

```{r}
maryland_ppp %>% 
  filter(PNHA==0, race=="Asian") %>% 
  group_by(zip5) %>% 
  summarise(total_amount=sum(amount),n=n())
  

```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?
**A7** There are 52 industries with exactly one loan in the data. I would like to do further research into who owns the Cotton Gin in Maryland and what are they using it for that they are requiring PPP loans. I think it will make for an interesting story about who is cotton ginning, why and how they still survive in the 21st century. 

```{r}
maryland_ppp %>% 
  group_by(naics_code, title) %>% 
  summarise(n=n()) %>% 
  filter(n==1)
```
